---
title: "HW4"
author: "Atharv Vani"
date: "2025-02-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(tidyverse)
library(mosaic)
```


## Name: Atharv Vani
## UT EID: amv4278
## Github Link: https://github.com/atharvvani464/SDS315-HW4


## Problem 1

```{r, dev = 'pdf'}
n_trades <- 2021
p_flag <- 0.024
observed_flags <- 70
n_simulations <- 100000

simulated_flags <- rbinom(n_simulations, size = n_trades, prob = p_flag)
p_value <- mean(simulated_flags >= observed_flags)

ggplot(data.frame(simulated_flags), aes(x = simulated_flags)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = observed_flags, linetype = "dashed") +
  labs(title = "Monte Carlo Distribution of Flagged Trades Under Null Hypothesis",
       x = "Number of Flagged Trades", y = "Frequency")

print(paste("P-Value:", p_value))

if (p_value < 0.05) {
  print("The p-value is low, suggesting that the number of flagged trades may be higher than expected under the null hypothesis.")
} else {
  print("The p-value is not sufficiently low to conclude that the number of flagged trades is unusually high.")
}
```
Null Hypothesis: The null hypothesis being tested is that the rate of flagged trades is equal to the baseline probability of 2.4%.
Test Statistic: The test statistic used is the number of flagged trades observed.
A Monte Carlo simulation is used to generate the probability distribution of flagged trades under the null hypothesis. The p-value is computed to assess whether the observed number of flagged trades is significantly different from expectation.

## Problem 2

```{r, dev = 'pdf'}
total_inspections <- 1500
p_violation <- 0.03
n_gourmet_inspections <- 50
observed_violations <- 8
n_simulations <- 100000

simulated_violations <- rbinom(n_simulations, size = n_gourmet_inspections, prob = p_violation)

p_value <- mean(simulated_violations >= observed_violations)

ggplot(data.frame(simulated_violations), aes(x = simulated_violations)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = observed_violations, linetype = "dashed") +
  labs(title = "Monte Carlo Distribution of Violations Under Null Hypothesis",
       x = "Number of Violations", y = "Frequency")

print(paste("P-Value:", p_value))

if (p_value < 0.05) {
  print("The p-value is low, suggesting that Gourmet Bites may have a higher violation rate than expected under the null hypothesis.")
} else {
  print("The p-value is not sufficiently low to conclude that Gourmet Bites has an unusually high violation rate.")
}
```


Null Hypothesis: The null hypothesis being tested is that Gourmet Bites has the same health code violation rate as the citywide average of 3%.
Test Statistic: The test statistic used is the number of health code violations observed at Gourmet Bites during inspections.
A Monte Carlo simulation is performed to estimate the probability distribution of this test statistic under the null hypothesis. The p-value is computed to determine whether the observed violation rate is significantly different from the expected rate.

## Problem 3
```{r, dev = 'pdf'}
observed_counts <- c(85, 56, 59, 27, 13)
total_jurors <- sum(observed_counts)
expected_proportions <- c(0.30, 0.25, 0.20, 0.15, 0.10)
expected_counts <- total_jurors * expected_proportions
chi_squared_statistic <- function(observed, expected) {
  sum((observed - expected)^2 / expected)
}
num_simulations <- 10000
chi2_sim <- do(num_simulations) * {
  simulated_counts <- rmultinom(1, total_jurors, expected_proportions)
  this_chi2 <- chi_squared_statistic(simulated_counts, expected_counts)
  c(chi2 = this_chi2)
}

ggplot(chi2_sim) + 
  geom_histogram(aes(x = chi2), bins = 30)

my_chi2 <- chi_squared_statistic(observed_counts, expected_counts)

p_value <- mean(chi2_sim$chi2 >= my_chi2)
print(paste("P-Value:", p_value))

if (p_value < 0.05) {
  print("The p-value is low, suggesting that the jury selection process overseen by this judge significantly deviates from the county’s expected demographic distribution.")
} else {
  print("The p-value is not sufficiently low to conclude that the jury selection process is biased.")
}
```


